---
title: "Hyperoptic internet speed"
project:
  type: website
---

The past week or so has been full of troubleshooting wifi speed problems. Various changes suggested by the Hyperoptic team has not succeeded in bringing my internet speed close to the speed of 1 Gbps promised in the package I chose. 

Most recent suggestion was that I try doing the speedtest on different devices. Perhaps the experts at hyperoptic are baffled their suggestions of trying a different cable or resetting the router are not working. The trouble must be with my device! So this is what I did. 

However, I wondered, how will they know if the difference in speed between different devices is significant? Worry not, Hyperoptic, I was trained in statistics, and my passion for `R` can help us both. 

You are welcome. 

```{r}
#| output: false
#| echo: false
library(tidyverse)
library(car)
library(dplyr)
library(ggplot2)
library(reshape2)
library(report)
library(multcomp)

data <- read.csv("Speedtest_ResultsHistory.csv", header = TRUE)

# select relevant columns
data <- data %>%
  dplyr::select(Download, Upload, device)

# change units
data$Download <- as.numeric(data$Download) / 1000
data$Upload <- as.numeric(data$Upload) / 1000
data$device <- as.factor(data$device)
```

## Create a summary of data

```{r}
#| output: false
#| echo: true
summary <- group_by(data, device) %>%
  summarise(
    mean_download = mean(Download, na.rm = TRUE),
    sd_download = sd(Download, na.rm = TRUE),
    mean_upload = mean(Upload, na.rm = TRUE),
    sd_upload = sd(Upload, na.rm = TRUE)
)

print(summary)
```

The average speed per device are:

|  | download | upload |
|---|---|---|
| personal laptop | `{r} print(round(summary$mean_download[1], 0))` Mbps ±`{r} print(round(summary$sd_download[1], 1))` | `{r} print(round(summary$mean_upload[1], 0))` Mbps ± `{r} print(round(summary$sd_upload[1], 0))` |
| phone | `{r} print(round(summary$mean_download[2], 0))` Mbps ±`{r} print(round(summary$sd_download[2], 1))` | `{r} print(round(summary$mean_upload[2], 0))` Mbps ± `{r} print(round(summary$sd_upload[2], 0))` |
| work laptop | `{r} print(round(summary$mean_download[1], 0))` Mbps ±`{r} print(round(summary$sd_download[3], 1))` | `{r} print(round(summary$mean_upload[3], 0))` Mbps ±`{r} print(round(summary$sd_upload[3], 0))` |

Averages are calculated for 10 measurements per device. Standar deviation values are denoted by ±. 

## Visualise data

@boxplot is a visualisation of the values summarised above.

```{r}
#| label: boxplot
#| fig-cap: "Boxplot of download and upload values per device"
#| warning: false

data_melted <- melt(data, id = "device")

ggplot(data = data_melted) +
  geom_boxplot(aes(x = device, y = value, fill = variable)) +
  xlab("device") +
  ylab("speed (Mbps)") +
  labs(fill = "speed measured")
```

## Test the normality of data
Conduct Shapiro test to check normality of data. Non-significant p-values (lower than 0.05) indicate the data is normally distributed.

### Download speed:

```{r}
#| output: true
#| echo: true
res_download <- aov(Download ~ device, data = data)
normal_download <- shapiro.test(res_download$residuals)

print(normal_download)
```

The p-value of Shapiro test is `{r} print(round(normal_download$p.value, 3))`, meaning the data is normally distributed.

### Upload speed:

```{r}
#| output: true
#| echo: true
res_upload <- aov(Upload ~ device, data = data)
normal_upload <- shapiro.test(res_upload$residuals)

print(normal_upload)
```

The p-value of Shapiro test is `{r} print(round(normal_upload$p.value, 3))`, meaning the data is normally distributed.

## Check homoscedascidity of data
Conduct Levene's test to check for homoscedascidity of data.

### Download speed:
```{r}
#| output: true
#| echo: true
homoscedasticity_d <- leveneTest(Download ~ factor(device),
  data = data)

print(homoscedasticity_d)
```

Levene's test is significant, meaning variances of download speed data across different devices are not equal.

### Upload speed:
```{r}
homoscedasticity_u <- leveneTest(Upload ~ factor(device),
  data = data)

print(homoscedasticity_u)
```

Levene's test is significant, meaning our data is heteroscedastic (i.e. variances of upload speed data across different devices are not equal).

## ANOVA

### Assumptions
The assumption of:

- **Normality of data** has been met- both download and upload speed values are normally distributed.
- **Homoscedasticity** has **not** been met- download and upload speed values have unequal variances between types of device. 
- **Independence of samples** has been met.  

Our data is heteroscedastic, and therefore one of the assumptions of ANOVA has been violated, however at this point I don't have the determination to research bootstrapping methods to corect for this, so i will just proceed with the ANOVA.

### Analysis of variance (ANOVA)

#### Download ANOVA
```{r}
#| output: true
#| echo: true
print(report(res_download))
```

#### Upload ANOVA

```{r}
#| output: true
#| echo: true
print(report(res_upload))
```

### ANOVA results

There was a main significant effect of device type on both upload and download speed.

To investigate which devices were significantly different from each other I conducted a post hoc test.

### Turkey HSD test

#### Download Turkey

```{r}
#| output: true
#| echo: true
post_test_d <- glht(res_upload,
  linfct = mcp(device = "Tukey")
)

print((summary(post_test_d, test = univariate())))
```

#### Upload Turkey

```{r}
#| output: true
#| echo: true
post_test_u <- glht(res_upload,
  linfct = mcp(device = "Tukey")
)

print((summary(post_test_u, test = univariate())))
```

## Interpretation of results 

According to the resulrs of Turkey test:

- There is a significant difference in **download** speed between `phone and personal laptop` and between `phone and work laptop`. However, the comparison between `personal laptop and work laptop` is not significant.

- Similarily, there is a significant difference in **upload** speed between `phone and personal laptop` and between `phone and work laptop`. However, the comparison between `personal laptop and work laptop` is not significant.

